---
title: "DA_Pancreatic_DE"
output: html_document
date: "2023-10-05"
editor_options: 
  chunk_output_type: console
---
# Set up

## Load packages

```{r set up}
library(tidyverse)
library(limma)
library(pheatmap)
library(ggbeeswarm)
library(ggrepel)
library(ggplotify)
library(patchwork)
library(clusterProfiler)
library(enrichplot)
library(corrplot)

select <- dplyr::select

source("scripts/functions_analyses.R")
source("scripts/functions_utility.R")
source("scripts/themes_palettes.R")
source("scripts/functions_visualization.R")
```

## Disease Atlas data

```{r read DA data}
#| include: false
cancer_data_long <-  read_csv("data/20230629_cancer_data.csv")
cancer_metadata <- read_csv("data/20230629_cancer_metadata.csv")

# Make data wide
cancer_data_wide <-
  cancer_data_long |> 
  select(DAid, Disease, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX")

pancreatic_meta <- readxl::read_excel("data/clinical_metadata/Pancreatic/UPApatients.xlsx")
da_meta <- read_csv("data/processed/final_data/final_metadata_all_batches.csv") 
da_data <- read_csv("data/processed/final_data/final_data_all_batches.csv") 
dw_info <- readxl::read_excel("data/samples_2023-05-12.xlsx")
```

### Subsample DA data

```{r Subsample other cancers}
# Minimum is myeloma - 41 -> 40 from each disease
cancer_metadata |> 
  filter(!Disease %in% c("PAN", "CAD healthy")) |> 
  group_by(Disease) |> 
  summarise(n = n_distinct(DAid)) |> 
  arrange(n)

set.seed(213)
sampled_other_cancers <- 
  cancer_metadata |> 
  filter(!Disease %in% c("PAN", "CAD healthy")) |> 
  group_by(Disease) |> 
  sample_n(40)

final_meta <- 
  sampled_other_cancers |> 
  bind_rows(cancer_metadata |> 
              filter(Disease %in% c("PAN", "CAD healthy"))) |> 
  mutate(Disease = case_when(Disease == "PAN" ~ "Pancreatic cancer",
                             Disease == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) 
```

# Differential expression analyses

## Compared to healthy

```{r}
# Prepare data & metadata
patients_de <- 
  final_meta |> 
  filter(Disease %in% c("Pancreatic cancer", "Healthy")) |> # 941 -> 931 (not many exclusions)
  filter(!is.na(Sex),
         !Age < 10,
         !is.na(Age))

wide_data <- 
  cancer_data_wide |> 
  select(-Disease) |> 
  filter(DAid %in% patients_de$DAid) 

# Run differential expression analysis
de_healthy <- 
  de_limma_disease(data_wide = wide_data, 
                   metadata = patients_de,
                   disease = "Pancreatic cancer",
                   correct = T) |> 
  mutate(sig = case_when(adj.P.Val < 0.05 & logFC < -1 ~ "significant down",
                         adj.P.Val < 0.05 & logFC > 1 ~ "significant up", 
                         T ~ "not significant"))

# Generate volcano plot
volcano_healthy <- plot_volcano(de_healthy)
ggsave(savepath("de_pancreatic_DA.png"), h = 5, w = 5)
```

## MEDECA comparison

```{r}
de_medeca <- read_tsv("data/de_medeca.tsv")

joint_dat <- 
  de_healthy |> 
  select(Assay, logFC_DA = logFC, p_DA = adj.P.Val) |>
  left_join(de_medeca |> 
              filter(Cancer == "Pancreas or gall bladder or bile duct") |> 
              select(Assay, logFC_MEDECA = logFC, p_MEDECA = adj.P.Val), by = "Assay") 

top20_pval <- 
  de_healthy |> 
  top_n(30, -adj.P.Val)

top20_fc <- 
  de_healthy |> 
  top_n(30, logFC)

joint_dat |> 
  mutate(top30_FC_DA = ifelse(Assay %in% top20_fc$Assay, "Yes", "No")) |> 
  ggplot(aes(logFC_DA, logFC_MEDECA, color = top30_FC_DA)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  geom_hline(yintercept = 0, lty = "dashed", color = "grey") + 
  geom_vline(xintercept = 0, lty = "dashed", color = "grey") +
  scale_color_manual(values = c("grey", "darkred")) +
  theme_hpa()  +
  
joint_dat |> 
  mutate(top30_pval_DA = ifelse(Assay %in% top20_pval$Assay, "Yes", "No")) |> 
  ggplot(aes(-log10(p_DA), -log10(p_MEDECA), color = top30_pval_DA)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  geom_hline(yintercept = -log10(0.05), lty = "dashed", color = "grey") + 
  geom_vline(xintercept = -log10(0.05), lty = "dashed", color = "grey") +
   scale_color_manual(values = c("grey", "darkred")) +
  theme_hpa() 

ggsave(savepath("pancreatic_MEDECA_comparison.png"), h = 5, w = 12)
```

## Compared to healthy turkey

```{r}
# Prepare data & metadata
patients_de <- 
  da_meta |> 
  mutate(Sex = case_when(Gender == 0 ~ "Female",
                         Gender == 1 ~ "Male")) |> 
  select(DAid, Disease, Age, Sex, BMI) |> 
  filter(Disease %in% c("Pancreatic cancer", "Healthy Turkey")) #|> # 941 -> 931 (not many exclusions)
  #filter(!is.na(Sex),
   #      !Age < 10,
    #     !is.na(Age))

wide_data <- 
  da_data |> 
#  select(-Disease) |> 
  filter(DAid %in% patients_de$DAid) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(values_from = "NPX", names_from = "Assay")

# Run differential expression analysis
de_healthy_turkey<- 
  de_limma_disease(data_wide = wide_data, 
                   metadata = patients_de,
                   disease = "Pancreatic cancer",
                   correct = F) |> 
  mutate(sig = case_when(adj.P.Val < 0.05 & logFC < -1 ~ "significant down",
                         adj.P.Val < 0.05 & logFC > 1 ~ "significant up", 
                         T ~ "not significant"))

# Generate volcano plot
volcano_healthy <- plot_volcano(de_healthy_turkey)

de_healthy |> 
  select(Assay, logFC) |> 
  left_join(de_healthy_turkey |> 
              select(Assay, logFC_turkey = logFC)) |> 
  ggplot(aes(logFC, logFC_turkey)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  theme_hpa()

```


## Compared to Wellness

```{r}
# Prepare data & metadata
patients_de <- 
  da_meta |> 
  mutate(Sex = case_when(Gender == 0 ~ "Female",
                         Gender == 1 ~ "Male")) |> 
  select(DAid, Disease, Age, Sex, BMI) |> 
  filter(Disease %in% c("Pancreatic cancer", "Wellness")) #|> # 941 -> 931 (not many exclusions)
  #filter(!is.na(Sex),
   #      !Age < 10,
    #     !is.na(Age))

wide_data <- 
  da_data |> 
#  select(-Disease) |> 
  filter(DAid %in% patients_de$DAid) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(values_from = "NPX", names_from = "Assay")

# Run differential expression analysis
de_healthy_wellness <- 
  de_limma_disease(data_wide = wide_data, 
                   metadata = patients_de,
                   disease = "Pancreatic cancer",
                   correct = F) |> 
  mutate(sig = case_when(adj.P.Val < 0.05 & logFC < -1 ~ "significant down",
                         adj.P.Val < 0.05 & logFC > 1 ~ "significant up", 
                         T ~ "not significant"))

# Generate volcano plot
plot_volcano(de_healthy_wellness)

de_healthy |> 
  select(Assay, logFC) |> 
  left_join(de_healthy_wellness |> 
              select(Assay, logFC_wellness = logFC)) |> 
  ggplot(aes(logFC, logFC_wellness)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  theme_hpa()

```


## Compared to EpiHealth

```{r}
# Prepare data & metadata
patients_de <- 
  da_meta |> 
  mutate(Sex = case_when(Gender == 0 ~ "Female",
                         Gender == 1 ~ "Male")) |> 
  select(DAid, Disease, Age, Sex, BMI) |> 
  filter(Disease %in% c("Pancreatic cancer", "EpiHealth")) #|> # 941 -> 931 (not many exclusions)
  #filter(!is.na(Sex),
   #      !Age < 10,
    #     !is.na(Age))

wide_data <- 
  da_data |> 
#  select(-Disease) |> 
  filter(DAid %in% patients_de$DAid) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(values_from = "NPX", names_from = "Assay")

# Run differential expression analysis
de_healthy_epi <- 
  de_limma_disease(data_wide = wide_data, 
                   metadata = patients_de,
                   disease = "Pancreatic cancer",
                   correct = F) |> 
  mutate(sig = case_when(adj.P.Val < 0.05 & logFC < -1 ~ "significant down",
                         adj.P.Val < 0.05 & logFC > 1 ~ "significant up", 
                         T ~ "not significant"))

# Generate volcano plot
plot_volcano(de_healthy_epi)

de_healthy |> 
  select(Assay, logFC) |> 
  left_join(de_healthy_epi |> 
              select(Assay, logFC_epi = logFC)) |> 
  ggplot(aes(logFC, logFC_epi)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  theme_hpa()

plot_volcano(de_ukb)


de_healthy |> 
  select(Assay, logFC) |> 
  left_join(de_ukb |> 
              select(Assay, logFC_ukb = logFC)) |> 
  ggplot(aes(logFC, logFC_ukb)) +
  geom_point() +
  geom_text_repel(aes(label = Assay)) +
  theme_hpa()


```


## Compared to other cancers

```{r}
# Prepare data & metadata
patients_de_cancers <- 
  final_meta |> 
  filter(Disease %in% c("Pancreatic cancer", "Other cancers")) |> # 594 -> 569 (not many exclusions)
  filter(!is.na(Sex),
         !Age < 10,
         !is.na(Age)) 

wide_data_cancers <- 
  cancer_data_long |> 
  filter(DAid %in% patients_de_cancers$DAid) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX") 

# Run differential expression analysis
de_cancers <- 
  de_limma_disease(data_wide = wide_data_cancers, 
                   metadata = patients_de_cancers,
                   disease = "Pancreatic cancer",
                   correct = T)  |> 
  mutate(sig = case_when(adj.P.Val < 0.05 & logFC < -1 ~ "significant down",
                         adj.P.Val < 0.05 & logFC > 1 ~ "significant up", 
                         T ~ "not significant"))

# Generate volcano plot
volcano_cancers <- plot_volcano(de_cancers)
```


# Summary

## Comparison

```{r}
# Combine results
de_combined <- 
  de_healthy |> 
  select(Assay, logFC_healthy = logFC, adjpval_healthy = adj.P.Val, sig_healthy = sig) |> 
  left_join(de_cancers |> 
              select(Assay, logFC_cancers = logFC, adjpval_cancers = adj.P.Val, sig_cancers = sig), by = "Assay") |> 
  mutate(Significance = ifelse(sig_healthy == sig_cancers, sig_healthy, "not significant")) 


de_combined |> 
  saveRDS("data/processed/DE/DE_combined.rds")
# Compare log fold change
comparison_fc <- 
  de_combined |> 
  ggplot(aes(logFC_healthy, logFC_cancers, label = Assay, color = Significance)) +
  geom_point() +
  geom_text_repel(show.legend = F, size = 3) +
  scale_color_manual(values = pal_de) +
  theme_hpa() 

# Figure 2 A,C & C
#volcano_healthy + volcano_cancers + comparison_fc
#ggsave(savepath("volcanos.pdf"), h = 4, w = 12)
```

## Overlap 

```{r}

# Supplementary figure C

```

## Boxplots - top proteins 

```{r}
de_combined |> 
  filter(Significance != "not significant") |> 
  mutate(logFC_avg = (abs(logFC_cancers) + abs(logFC_healthy))/2) |> 
  group_by(Significance) |> 
  top_n(4, logFC_avg) |> 
  arrange(Significance,-logFC_avg)

top_up <- c("PTN", "MUC16", "FGF21", "CCL20")
top_down <-  c("GPA33", "SSC4D", "TMPRSS15", "TNFSF11")

long_levels <- 
  cancers_mapping |> 
  mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
         Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer),
         Cancer_code = factor(Cancer_code, levels = all_cancers_levels)) |> 
  arrange(Cancer_code) |> 
  pull(Cancer)

upregulated_prots <- 
  cancer_data_long |> 
  filter(Assay %in% top_up,
         Disease %in% all_levels) |> 
  rename(GROUP = Disease) |> 
  mutate(Disease = case_when(GROUP == "PAN" ~ "Pancreatic cancer",
                             GROUP == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) |> 
  left_join(cancers_mapping |> 
              mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
                     Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer)), 
            by = c("GROUP" = "Cancer_code")) |> 
  mutate(Cancer = ifelse(GROUP == "CAD healthy", "Healthy", Cancer),
         Assay = factor(Assay, levels = top_up),
         Cancer = factor(Cancer, levels = c(long_levels, "Healthy"))) |> 
  select(DAid, Cancer, Disease, Assay, NPX) |> 
  ggplot(aes(Cancer, NPX, color = Disease)) +
  geom_quasirandom(alpha = 0.4, show.legend = F) +
  geom_boxplot(fill = NA, color = "grey60", outlier.color = NA) +
  facet_wrap(~ Assay, ncol = 2, scales = "free_y") +  
  scale_color_manual(values = pal_group) +
  xlab("") +
  theme_hpa()


downregulated_prots <- 
  cancer_data_long |> 
  filter(Assay %in% top_down,
         Disease %in% all_levels) |> 
  rename(GROUP = Disease) |> 
  mutate(Disease = case_when(GROUP == "PAN" ~ "Pancreatic cancer",
                             GROUP == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) |> 
  left_join(cancers_mapping |> 
              mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
                     Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer)), 
            by = c("GROUP" = "Cancer_code")) |> 
  mutate(Cancer = ifelse(GROUP == "CAD healthy", "Healthy", Cancer),
         Assay = factor(Assay, levels = top_down),
         Cancer = factor(Cancer, levels = c(long_levels, "Healthy"))) |> 
  select(DAid, Cancer, Disease, Assay, NPX) |> 
  ggplot(aes(Cancer, NPX, color = Disease)) +
  geom_quasirandom(alpha = 0.4, show.legend = F) +
  geom_boxplot(fill = NA, color = "grey60", outlier.color = NA, show.legend = F,linewidth = 0.5) +
  facet_wrap(~ Assay, ncol = 2, scales = "free_y") +  
  scale_color_manual(values = pal_group) +
  xlab("") +
  theme_hpa()


upregulated_prots <- 
  cancer_data_long |> 
  filter(Assay %in% c(top_up),
         Disease %in% all_levels) |> 
  rename(GROUP = Disease) |> 
  mutate(Disease = case_when(GROUP == "PAN" ~ "Pancreatic cancer",
                             GROUP == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) |> 
  left_join(cancers_mapping |> 
              mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
                     Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer)), 
            by = c("GROUP" = "Cancer_code")) |> 
  mutate(Cancer = ifelse(GROUP == "CAD healthy", "Healthy", Cancer),
         Assay = factor(Assay, levels = c(top_up)),
         Cancer = factor(Cancer, levels = c("Healthy", "Pancreatic cancer", long_levels[-14]))) |> 
  select(DAid, Cancer, Disease, Assay, NPX) |> 
  ggplot(aes(Cancer, NPX, color = Disease, fill = Disease)) +
  geom_quasirandom(size = 0.8, alpha = 0.4, stroke = NA, show.legend = F) +
  geom_boxplot(alpha = 0.4, color = "grey30", outlier.color = NA, show.legend = F, linewidth = 0.5) +
  facet_wrap(~ Assay, nrow = 4, scales = "free_y") +  
  scale_color_manual(values = pal_group) +
  scale_fill_manual(values = pal_group) +
  #coord_flip() +
  xlab("") +
  theme_hpa(angled = T)

downregulated_prots <- 
  cancer_data_long |> 
  filter(Assay %in% c(top_down),
         Disease %in% all_levels) |> 
  rename(GROUP = Disease) |> 
  mutate(Disease = case_when(GROUP == "PAN" ~ "Pancreatic cancer",
                             GROUP == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) |> 
  left_join(cancers_mapping |> 
              mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
                     Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer)), 
            by = c("GROUP" = "Cancer_code")) |> 
  mutate(Cancer = ifelse(GROUP == "CAD healthy", "Healthy", Cancer),
         Assay = factor(Assay, levels = c(top_down)),
         Cancer = factor(Cancer, levels = c("Healthy", "Pancreatic cancer", long_levels[-14]))) |> 
  select(DAid, Cancer, Disease, Assay, NPX) |> 
  ggplot(aes(Cancer, NPX, color = Disease, fill = Disease)) +
  geom_quasirandom(size = 0.8, alpha = 0.4, stroke = NA, show.legend = F) +
  geom_boxplot(alpha = 0.4, color = "grey30", outlier.color = NA, show.legend = F, linewidth = 0.5) +
  facet_wrap(~ Assay, nrow = 4, scales = "free_y") +  
  scale_color_manual(values = pal_group) +
  scale_fill_manual(values = pal_group) +
  #coord_flip() +
  xlab("") +
  theme_hpa(angled = T)


upregulated_prots + downregulated_prots

ggsave(savepath("boxplots_de.pdf"), h = 7, w = 6)
ggsave(savepath("boxplots_de.pdf"), h = 7, w = 8)

# Alternative
cancer_data_long |> 
  filter(Assay %in% c(top_up, top_down),
         Disease %in% all_levels) |> 
  rename(GROUP = Disease) |> 
  mutate(Disease = case_when(GROUP == "PAN" ~ "Pancreatic cancer",
                             GROUP == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) |> 
  left_join(cancers_mapping |> 
              mutate(Cancer = ifelse(Cancer_code %in% c("AML", "CLL"), Cancer_code, Cancer),
                     Cancer = ifelse(Cancer_code %in% c("LYMPH"), "DLBCL", Cancer)), 
            by = c("GROUP" = "Cancer_code")) |> 
  mutate(Cancer = ifelse(GROUP == "CAD healthy", "Healthy", Cancer),
         Assay = factor(Assay, levels = c(top_up, top_down)),
         Cancer = factor(Cancer, levels = c(long_levels, "Healthy"))) |> 
  select(DAid, Cancer, Disease, Assay, NPX) |> 
  ggplot(aes(Cancer, NPX, color = Disease)) +
  geom_quasirandom(alpha = 0.4, show.legend = F) +
  geom_boxplot(fill = NA, color = "grey60", outlier.color = NA) +
  facet_wrap(~ Assay, ncol = 8, scales = "free_x") +  
  scale_color_manual(values = pal_group) +
  coord_flip() +
  xlab("") +
  theme_hpa()

ggsave(savepath("boxplots_de.pdf"), h = 4, w = 16)

# Figure 2 D, E

```

## Heatmap - top proteins 


```{r}
significant_prots <-
  de_combined |>
  filter(Significance != "not significant") |>
  arrange(Significance, logFC_cancers) |>
  select(Assay, Significance) |>
  left_join(hpa_info_short |> 
              select(Assay = Gene, Secretome = `Secretome location`) |> 
              mutate(Secretome = ifelse(is.na(Secretome), "Unknown", Secretome)), by = "Assay") |> 
  column_to_rownames("Assay")

heatmap_samples <- 
  final_meta |> 
  filter(Disease != "Other cancers") |> 
  select(DAid, Disease) |> 
  column_to_rownames("DAid")

cancer_data_long |> 
  filter(Assay %in% rownames(significant_prots),
         DAid %in% rownames(heatmap_samples)) |> 
  mutate(Assay = factor(Assay, levels = rownames(significant_prots)),
         DAid = factor(DAid, levels = rownames(heatmap_samples))) |> 
  filter(Disease %in% c("CAD healthy", "PAN")) |> 
  select(DAid, Assay, NPX) |> 
  group_by(Assay) |> 
  mutate(NPX = scales::rescale(NPX, to = c(0,1))) |> 
  arrange(DAid, Assay) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX") |> 
  column_to_rownames("DAid") |> 
  pheatmap(show_rownames = F,
           clustering_method = "ward.D2",
           # cluster_rows = F, 
           #cluster_cols = F,
           border_color = NA,
           annotation_col =  significant_prots,
           annotation_row = heatmap_samples,
           annotation_colors = list(Disease = pal_group,
                                    Significance = pal_de[-1])) |> 
  as.ggplot()


ggsave(savepath("heatmap_de.pdf"), h = 4, w = 14)

library(RColorBrewer)
rd_bu_continuous <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))

cancer_data_long |> 
  filter(Assay %in% rownames(significant_prots),
         DAid %in% rownames(heatmap_samples)) |> 
  mutate(Assay = factor(Assay, levels = rownames(significant_prots)),
         DAid = factor(DAid, levels = rownames(heatmap_samples))) |> 
  filter(Disease %in% c("CAD healthy", "PAN")) |> 
  select(DAid, Assay, NPX) |> 

  pivot_wider(names_from = "Assay", values_from = "NPX") |> 
  column_to_rownames("DAid") |> 
  scale() |> 
  pheatmap(show_rownames = F,
           color = rd_bu_continuous,
           clustering_method = "ward.D2",

          border_color = NA,
           annotation_col =  significant_prots,
           annotation_row = heatmap_samples,
           annotation_colors = list(Disease = pal_group,
                                    Secretome = pal_secreted,
                                    Significance = pal_de[-1])) |> 
  as.ggplot()

ggsave(savepath("healtmap_DE.pdf"), h = 6, w = 16)

significant_prots |> 
  rownames_to_column("Gene") |> 
  left_join(hpa_info_short |> 
              select(Gene, secretome = `Secretome location`))
```

## Correlation

### All DE proteins

```{r}
de_prots <-
  de_combined |>
  filter(Significance != "not significant") |>
  select(-sig_healthy, -sig_cancers) |>
  mutate(log10pval_healthy = -log10(adjpval_healthy),
         log10pval_cancers= -log10(adjpval_cancers)) |> 
  select(Assay, logFC_healthy, logFC_cancers, log10pval_healthy, log10pval_cancers, Significance) |> 
  column_to_rownames("Assay")

corr_mat <- 
  cancer_data_long |> 
  filter(Assay %in% rownames(significant_prots),
         DAid %in% rownames(heatmap_samples)) |> 
  mutate(Assay = factor(Assay, levels = rownames(significant_prots)),
         DAid = factor(DAid, levels = rownames(heatmap_samples))) |> 
  filter(Disease %in% c("CAD healthy", "PAN")) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX") |> 
  column_to_rownames("DAid") |> 
  scale() |> 
  cor(use = "complete.obs")

colors <- 
  de_prots |> 
  group_by_all() |> 
  mutate(color = pal_de[[Significance]]) |> 
  ungroup() |> 
  pull(color)

ord <- corrMatOrder(corr_mat, order="hclust", hclust.method = "ward.D2")

new_colours <- 
  colnames(corr_mat)[ord] |> 
  enframe() |> 
  left_join(de_prots |> 
              rownames_to_column("value"), by = "value") |> 
  group_by_all() |> 
  mutate(color = pal_de[[Significance]]) |> 
  ungroup() |> 
  select(value, color) |> 
  deframe()

corrplot(corr_mat, 
         method = "square",
         type="lower", 
         order="hclust", 
         hclust.method = "ward.D2",
         col=pal_corr,  
         tl.srt=0,
         tl.col = new_colours) |> 
  as.ggplot()

# Supplememntary Figure 4A
ggsave(savepath("corr_de.pdf"), h = 10, w = 10)

corr_mat |> 
  pheatmap(annotation_row = de_prots,
           color = col(200),
           annotation_colors = list(Significance = pal_de[-1]))
```

### High correlation examples

```{r}
corr_mat |> 
  as_tibble(rownames = "Protein1") |> 
  pivot_longer(cols = -Protein1,
               names_to = "Protein2", 
               values_to = "cor") |> 
  filter(Protein1 != Protein2) |> 
  arrange(-abs(cor))

dat <- 
  cancer_data_long |> 
  filter(DAid %in% rownames(heatmap_samples)) |> 
  mutate(DAid = factor(DAid, levels = rownames(heatmap_samples))) |> 
  filter(Disease %in% c("CAD healthy", "PAN")) |> 
  select(DAid, Disease,Assay, NPX) |> 
  pivot_wider(names_from = "Assay", 
              values_from = "NPX") 

dat |> 
  mutate(Disease = recode(Disease,
                           `CAD healthy` = "Healthy",
                           PAN = "Pancreatic cancer")) |> 
  ggplot(aes(REG1B, REG1A, color = Disease)) +
  geom_point() +
  geom_text(aes(y = 5, x = 0, 
                label = paste0("Pearson's r: ", round(cor(REG1B, REG1A),2))),
            inherit.aes = F,
            color = "black") +
  scale_color_manual(values = pal_group) +
  theme_hpa()

# Supplementary Figure 4 C
ggsave(savepath("cor_REG1A_REG1B.pdf"), h = 4, w = 6)

dat |> 
  mutate(Disease = recode(Disease,
                           `CAD healthy` = "Healthy",
                           PAN = "Pancreatic cancer")) |> 
  ggplot(aes(EPCAM, GPA33, color = Disease)) +
  geom_point() +
   geom_text(aes(y = 3, x = -2, 
                label = paste0("Pearson's r: ", round(cor(EPCAM, GPA33),2))),
            inherit.aes = F,
            color = "black") +
  scale_color_manual(values = pal_group) +
  theme_hpa()

# Supplementary Figure 4 C
ggsave(savepath("cor_EPCAM_GPA33.pdf"), h = 4, w = 6)


# REG1B    REG1A 
# EPCAM    GPA33
# AGXT     HAO1    
# PTN      SFRP1
# CSF1     VSIG4
# PTN       MDK 
```

### Correlation distribution

```{r}
corr_mat_all <- 
  cancer_data_long |> 
  filter(DAid %in% rownames(heatmap_samples)) |> 
  filter(Disease %in% c("CAD healthy", "PAN")) |> 
  select(DAid, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX") |> 
  column_to_rownames("DAid") |> 
  scale() |> 
  cor(use = "complete.obs")
 
df_corr_all <- 
  corr_mat_all |> 
  as_tibble(rownames = "Protein1") |> 
  pivot_longer(cols = -Protein1,
               names_to = "Protein2", 
               values_to = "cor") |> 
  filter(Protein1 != Protein2) |> 
  left_join(de_combined |> 
              mutate(Significance = ifelse(Significance == "not significant", "s1", Significance)) |> 
              select(Protein1 = Assay, s1 = Significance), by = "Protein1") |> 
  left_join(de_combined |> 
              mutate(Significance = ifelse(Significance == "not significant", "s2", Significance)) |> 
              select(Protein2 = Assay, s2 = Significance), by = "Protein2") |> 
  mutate(type = ifelse(s1 == s2, "intra", "inter")) 

mu <- 
  df_corr_all |> 
  group_by(type) |> 
  summarise(cor = median(cor))

df_corr_all |> 
  ggplot(aes(cor, fill = type, color = type)) +
  geom_density(alpha = 0.6, color = NA) +
  geom_vline(data = mu, aes(xintercept = cor, color = type),
             show.legend = F, linetype = "dashed") +
  scale_color_manual(values = c("grey80", "grey20")) +
  scale_fill_manual(values = c("grey80", "grey20")) +
  theme_hpa()

# Supplementary figure 4 B
ggsave(savepath("cor_density.pdf"), h = 4, w = 7)
```



# Enrichment - GSEA

```{r}
original_gene_list <- de_combined$logFC_healthy
names(original_gene_list) <- 
  de_combined |> 
  left_join(cancer_data_long |> 
              distinct(Assay, UniProt), by = "Assay") |> 
  pull(UniProt)

gene_list = sort(original_gene_list, decreasing = TRUE)
```


## KEGG

```{r}
kegg_organism <-  "hsa"

kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               #nPerm        = 10000,
               #minGSSize    = 3,
               #maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "BH",
               keyType       = "uniprot")

# Dotplot
dotplot(kk2, showCategory = 20, title = "Enriched Pathways" , split=".sign") + facet_grid(.~.sign) + theme_hpa()

ggsave(savepath("gsea_kegg.pdf"), h = 6, w = 6)

# Pathway network - Figure 2 F
cnetplot(edox, categorySize="pvalue", color.params = list(foldChange = gene_list))
ggsave(savepath("KEGG_GSEA.pdf"), h = 10, w = 10)
```

### Pathways

```{r}
dme <- pathview(gene.data = kegg_gene_list, 
                pathway.id = "hsa05212", 
                species = kegg_organism)
```

### IL-17 pathway proteins

```{r}
# cancer_data_long |> 
#   distinct(Assay) |> 
#   filter(grepl("DEF", Assay))

proteins <-   c("CXCL8", "CCL20", "IL17F", "IL17C", "IL17RA", "IL17A", "TRAF2", "CASP8", "IL1B", "TNF", "IL6",
                "MMP9", "LCN2", "CXCL8", "CCL20", "CCL7", "CCL2", "CSF3", "DEFA1_DEFA1B", "DEFB4A_DEFB4B")
 cancer_data_long |>
  filter(Assay %in% proteins, 
         Disease %in% c("PAN", "CAD healthy")) |>
  mutate(Disease = recode(Disease, 
                          PAN = "Pancreatic cancer",
                          `CAD healthy` = "Healthy")) |> 
  select(Sample = DAid, Assay, NPX, Disease) |>
  mutate(Disease = factor(Disease, levels = names(pal_group))) |> 
  ggplot(aes(Disease,NPX, color = Disease, fill = Disease)) +
  geom_quasirandom(alpha = 0.8) +
  geom_boxplot(alpha = 0.2, color = "black", outlier.colour = NA) +
  facet_wrap(~Assay, scales = "free_y", nrow = 2) +
  theme_hpa(angled = T) +
  scale_color_manual(values = c(pal_group, pal_ukb)) +
  scale_fill_manual(values = c(pal_group, pal_ukb)) +
  xlab("") +
  ylab("")
 
ggsave(savepath("il17_examples.pdf"), h = 6, w = 15)
```


## GO BP

```{r}
gse <- gseGO(geneList = gene_list, 
             ont = "BP", 
             keyType = "UNIPROT", 
             #  nPerm = 10000, 
             #  minGSSize = 3, 
             #  maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = "org.Hs.eg.db", 
             pAdjustMethod = "BH")

# Dotplot
dotplot(gse, showCategory=50, split=".sign") + facet_grid(.~.sign) + theme_hpa()

# Pathway network
gse_translated <- setReadable(gse, 'org.Hs.eg.db', 'UNIPROT')
cnetplot(gse_translated, categorySize="pvalue", foldChange=gene_list, showCategory = 5)

ggsave(savepath("GO_BP_top5.png"), h = 10, w = 10)
```


# Citations

```{r}
packageVersion("limma")
packageVersion("clusterProfiler")

```

