---
title: "03_machine_learning"
output: html_document
date: "2024-10-31"
editor_options: 
  chunk_output_type: console
---

# Set up

```{r}
library(HDAnalyzeR)

source("scripts/functions/functions_analyses.R")
source("scripts/functions/functions_utility.R")
source("scripts/functions/functions_visualization.R")
source("scripts/functions/themes_palettes.R")


# Read in data (pan-disease)
data_disease <- import_df("../Human-disease-blood-atlas/data/final_data/HPA/v24_2/disease_data_phase1.tsv")
meta_disease <- import_df("../Human-disease-blood-atlas/data/final_data/HPA/v24_2/disease_meta_phase1.tsv")

# Pancreatic cancer metadata
pancreatic_meta <- readxl::read_excel("data/clinical_metadata/Pancreatic/UPApatients.xlsx")
```

### Subsample DA data

```{r Subsample other cancers}
# Minimum is myeloma - 41 -> 40 from each disease
cancer_metadata |> 
  filter(!Disease %in% c("PAN", "CAD healthy")) |> 
  group_by(Disease) |> 
  summarise(n = n_distinct(DAid)) |> 
  arrange(n)

set.seed(213)
sampled_other_cancers <- 
  cancer_metadata |> 
  filter(!Disease %in% c("PAN", "CAD healthy")) |> 
  group_by(Disease) |> 
  sample_n(40)

final_meta <- 
  sampled_other_cancers |> 
  bind_rows(cancer_metadata |> 
              filter(Disease %in% c("PAN", "CAD healthy"))) |> 
  mutate(Disease = case_when(Disease == "PAN" ~ "Pancreatic cancer",
                             Disease == "CAD healthy" ~ "Healthy",
                             T ~ "Other cancers")) 
```


# Data splitting

```{r}
ml_meta <-
  final_meta |> 
  filter(Disease %in% c("Pancreatic cancer", "Healthy"))

current_wide_data <- 
  cancer_data_long |>
  filter(Disease %in% c("PAN", "CAD healthy")) |> 
  mutate(Disease = recode(Disease, 
                          `CAD healthy` = "Healthy",
                          PAN = "Pancreatic cancer")) |> 
  select(DAid, Disease, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX") 

set.seed(213)
cancer_split <-
  current_wide_data |> 
  initial_split(prop = 0.7, strata = Disease)

cancer_train <- training(cancer_split)
cancer_test <- testing(cancer_split)

saveRDS(cancer_train, "data/processed/ML/cancer_train.rds")
saveRDS(cancer_test, "data/processed/ML/cancer_test.rds")

```

## Define general recipe

```{r}
cancer_recipe <- 
  recipe(Disease ~ ., data = cancer_train) |> 
  update_role(DAid, new_role = "id") |> 
  step_normalize(all_numeric()) |> 
  step_nzv(all_numeric()) |> 
  step_corr(all_numeric()) |> 
  step_select_roc(all_numeric(), outcome = "Disease", threshold = 0.7) |>  
  step_impute_knn(all_numeric()) |> 
  step_downsample(Disease)
```

### Check variables retained

```{r}
# ###
# train_preped <- prep(standardized) %>%
#   bake(new_data = NULL)
# 
# test_preped <-  prep(standardized) %>%
#   bake(new_data = test_data)
# 
# ?control_stack_grid()
# 
# library(vip)
# Check retained variables
#dat <- prep(cancer_recipe, cancer_train)
#dat$template |> dim()
#training_bake <- bake(cancer_recipe)
```

## Resamples

```{r}
# Generate resamples
set.seed(213)
cancer_rs <- vfold_cv(cancer_train, v = 10, strata = Disease)

# Define evaluation metrics for all workflows
eval_metrics <- metric_set(roc_auc, accuracy, sensitivity, specificity)

# Define control grid
#set.seed(213)
ctrl <- control_grid(save_pred = TRUE, 
                     parallel_over = "everything") 
ctrl <- control_grid(verbose = TRUE, 
                     allow_par = TRUE,
                     save_pred = TRUE, 
                     parallel_over = "everything") 
```

## Try random


```{r}
cancer_recipe_random <- 
  recipe(Disease ~ ., data = cancer_train) |> 
  update_role(DAid, new_role = "id") |> 
  step_normalize(all_numeric()) |> 
  step_nzv(all_numeric()) |> 
  step_corr(all_numeric()) |> 
  step_select_roc(all_numeric(), outcome = "Disease", threshold = 0.7) |>  
  step_impute_knn(all_numeric()) |> 
  step_shuffle(all_numeric()) |> 
  step_downsample(Disease)

glmnet_lasso_specs <-
  logistic_reg() |>
  set_mode("classification") |>
  set_engine("glmnet") |>
  set_args(penalty = tune(),
           mixture = 1) 

glmnet_wflow_random <-
  workflow() |> 
  add_recipe(cancer_recipe_random) |> 
  add_model(glmnet_lasso_specs) 

glmnet_wflow<-
  workflow() |> 
  add_recipe(cancer_recipe) |> 
  add_model(glmnet_lasso_specs) 

set.seed(213)
glmnet_grid <-
  glmnet_wflow_random |>
  extract_parameter_set_dials() |>
  grid_latin_hypercube(size = 5)

set.seed(213)
cancer_rs <- vfold_cv(cancer_train, v = 10, strata = Disease)

eval_metrics <- metric_set(roc_auc, accuracy, sensitivity, specificity)

set.seed(213)
ctrl <- control_grid(save_pred = TRUE, 
                     parallel_over = "everything"#, 
                     #extract = get_coefs
                     ) # look at extract = identity


set.seed(213)
glmnet_res_da_random <-
  glmnet_wflow_random |>
  tune_grid(
    resamples = cancer_rs,
    grid = glmnet_grid,
    control = ctrl,
    metrics = eval_metrics)

glmnet_res_da_random |> autoplot()

set.seed(213)
glmnet_res_da <-
  glmnet_wflow |>
  tune_grid(
    resamples = cancer_rs,
    grid = glmnet_grid,
    control = ctrl,
    metrics = eval_metrics)


glmnet_res_da |> autoplot()

best_lasso <- 
  glmnet_res_da |> 
  select_best("roc_auc")

final_glmnet_wf <- 
  glmnet_wflow |> 
  finalize_workflow(best_lasso)


final_lasso <- 
  fit(final_glmnet_wf, cancer_train)

final_lasso |> 
  extract_fit_parsnip() |> 
  tidy() |> 
  filter(abs(estimate)>0,
         term!= "(Intercept)") |> 
  arrange(-abs(estimate))

autoplot(glmnet_res_da)
```


# lasso

## Define specifications

```{r}
glmnet_lasso_specs <-
  logistic_reg() |>
  set_mode("classification") |>
  set_engine("glmnet") |>
  set_args(penalty = tune(),
           mixture = 1) 

# Set up lasso workflow
glmnet_wflow <-
  workflow() |> 
  add_recipe(cancer_recipe) |> 
  add_model(glmnet_lasso_specs) 
```

## Hyperparameter optimization

```{r}
# Define hyperparameter tuning grid
set.seed(213)
glmnet_grid <-
  glmnet_wflow |>
  extract_parameter_set_dials() |>
  grid_latin_hypercube(size = 30)

# Hyperparameter tuning
set.seed(213)
glmnet_res_da <-
  glmnet_wflow |>
  tune_grid(
    resamples = cancer_rs,
    grid = glmnet_grid,
    control = ctrl,
    metrics = eval_metrics)

# saveRDS(glmnet_res_da, "data/processed/ML_res/lasso_20230925.rds")
# glmnet_res_da <- readRDS("data/processed/ML_res/lasso_20230925.rds")

# Explore results and select the best performing hyperparameter combination
autoplot(glmnet_res_da)

best_glmnet <- 
  glmnet_res_da |> 
  select_best("roc_auc")
```

## Final fit

```{r}
final_glmnet <- 
  glmnet_wflow |> 
  finalize_workflow(best_glmnet)

final_glmnet_fit <- 
  last_fit(final_glmnet, cancer_split)
```

## Exploration


```{r}
# Explore performance
final_glmnet_fit |> 
  collect_metrics() 

final_glmnet_fit |> 
  collect_predictions() |> 
  mutate(Disease = factor(Disease, levels = c("Pancreatic cancer", "Healthy"))) |> 
  roc_auc(Disease, `.pred_Pancreatic cancer`) 

# Explore important proteins
important_proteins_da <- 
  final_glmnet_fit  |> 
  extract_fit_parsnip() %>%
  tidy() |> 
  filter(term != "(Intercept)") |> 
  arrange(-abs(estimate)) |> 
  filter(abs(estimate) > 0)

#saveRDS(important_proteins_da, "data/processed/pancreatic/da_lasso_importance.rds")

# Visualize important proteins
important_proteins_da |> 
  mutate(sign = ifelse(estimate > 0, "positive", "negative")) |> 
  ggplot(aes(abs(estimate), fct_reorder(term, abs(estimate)), fill = sign)) +
  geom_col(alpha  = 0.8) +
  theme_hpa() +
  scale_fill_manual(values = c("#FE625E", "#75B2E9")) +
  xlab("Model estimate") +
  ylab("") +
  ggtitle("lasso features")

# Figure 3 C
#ggsave(savepath("lasso_top_deatures.pdf"), h = 5, w = 5)

# ROC cruve

```



# random forest

## Define specifications

```{r}
cores <- parallel::detectCores()

rf_specs <-
  rand_forest() |>
  set_mode("classification") |>
  set_engine("ranger", importance = "impurity", num.threads = cores) |>
  set_args(mtry = tune(), min_n = tune(), trees = 1000) 

# Set up random forest workflow
rf_wflow <-
  workflow() |> 
  add_recipe(cancer_recipe) |> 
  add_model(rf_specs) 
```


## Hyperparameter optimization

```{r}
# Define hyperparameter tuning grid
set.seed(213) #test
rf_grid <-
  rf_wflow |>
  extract_parameter_set_dials() |>
  update(mtry = mtry(c(0, 40))) |>
  grid_latin_hypercube(size = 15)

# Hyperparameter tuning
set.seed(213)
rf_res_da <-
  rf_wflow |>
  tune_grid(
    resamples = cancer_rs,
    grid = rf_grid,
    control = ctrl,
    metrics = eval_metrics)

saveRDS(rf_res_da, "data/processed/ML_res/rf_20231006.rds")
#rf_res_da <- readRDS("data/processed/ML_res/rf_20231005.rds")

# Explore results and select the best performing hyperparameter combination
autoplot(rf_res_da)

best_rf <- 
  select_best(rf_res_da, metric = "roc_auc") |> 
  select(-.config)
```

## Final fit

```{r}
rf_wflow <- 
  rf_wflow |>  
  finalize_workflow(best_rf)

final_rf_fit <- last_fit(rf_wflow, cancer_split, metrics = eval_metrics) 
```

## Exploration

```{r}
# Explore performance
final_rf_fit |> 
  collect_metrics() 

final_rf_fit |> 
  collect_predictions() |> 
  mutate(Disease = factor(Disease, levels = c("Pancreatic cancer", "Healthy"))) |> 
  roc_auc(Disease, `.pred_Pancreatic cancer`) 

# Explore important proteins
important_proteins_rf <- 
  final_rf_fit  |> 
  extract_fit_parsnip() |> 
  vip::vi() |> 
  arrange(-Importance) |> 
  filter(Importance > 0)

#saveRDS(important_proteins_rf, "data/processed/pancreatic/da_rf_importance.rds")

# Visualize important proteins
important_proteins_rf |> 
  arrange(-Importance) |> 
  head(24) |> 
  ggplot(aes(Importance, fct_reorder(Variable, Importance))) +
  geom_col(alpha  = 0.8) +
  theme_hpa() +
  xlab("Importance") +
  ylab("") +
  ggtitle("random forest features")

important_proteins_rf |> 
  arrange(-Importance) |> 
  head(25) |> 
  filter(Variable %in% important_proteins_da$term)
# Figure 3C
ggsave(savepath("rf_top_deatures.pdf"), h = 5, w = 4)

# ROC cruve

```




# Comparison lasso / rf

## Overlap

```{r}
top_overlapping_prots <- 
  important_proteins_rf |> 
  head(24) |> 
  filter(Variable %in% important_proteins_da$term)

proteins_rf <- important_proteins_rf$Variable
proteins_lasso <- important_proteins_da$term

y <- list("RF" = proteins_rf, 
          "lasso" = proteins_lasso)
plot(euler(y, shape = "ellipse"), quantities = TRUE, fills = c("#D6E9C3", "#D5CDE1")) |> as.ggplot()

# Figure 3B
ggsave(savepath("overlap_ML.pdf"), h = 3, w = 3)

## Plot overlapping proteins 
cancer_data_long |>
  filter(Assay %in% top_overlapping_prots$Variable, 
         Disease %in% c("PAN", "CAD healthy")) |>
  mutate(Disease = recode(Disease, 
                          PAN = "Pancreatic cancer",
                          `CAD healthy` = "Healthy"),
         Type = ifelse(Assay %in% c("CSF1", "LTA4H", "PTN"), "Up", "Down")) |> 
  select(Sample = DAid, Type, Assay, NPX, Disease) |>
  mutate(Disease = factor(Disease, levels = names(pal_group)),
         Type = factor(Type, levels = c("Up", "Down"))) |> 
  ggplot(aes(Disease,NPX, color = Disease, fill = Disease)) +
  geom_quasirandom(alpha = 0.8) +
  geom_boxplot(alpha = 0.2, color = "black", outlier.colour = NA) +
  facet_wrap(Type~Assay, scales = "free_y", nrow = 1) +
  theme_hpa(angled = T) +
  scale_color_manual(values = c(pal_group)) +
  scale_fill_manual(values = c(pal_group)) +
  xlab("") +
  ylab("")

# Figure 3D
ggsave(savepath("examples_ml_2.pdf"), h = 3, w = 14)
#ggsave(savepath("examples_ml.pdf"), h = 6.5, w = 8)



cancer_data_long |>
  filter(Assay %in% top_overlapping_prots$Variable, 
         Disease %in% c("PAN", "CAD healthy")) |>
  mutate(Disease = recode(Disease, 
                          PAN = "Pancreatic cancer",
                          `CAD healthy` = "Healthy"),
         Type = ifelse(Assay %in% c("CSF1", "LTA4H", "PTN"), "Up", "Down")) |> 
  select(Sample = DAid, Type, Assay, NPX, Disease) |>
  mutate(Disease = factor(Disease, levels = names(pal_group)),
         Type = factor(Type, levels = c("Up", "Down"))) |> 
  ggplot(aes(Disease,NPX, color = Disease, fill = Disease)) +
  geom_quasirandom(alpha = 0.8) +
  geom_boxplot(alpha = 0.2, color = "black", outlier.colour = NA) +
  facet_wrap(Type~Assay, scales = "free_x", nrow = 1) +
  theme_hpa(angled = T) +
  coord_flip() +
  scale_color_manual(values = c(pal_group)) +
  scale_fill_manual(values = c(pal_group)) +
  xlab("") +
  ylab("")

# Figure 3D
ggsave(savepath("examples_ml.pdf"), h = 4, w = 12)
```

## Correlation

```{r}
corr_mat <- 
  cancer_data_wide |> 
  filter(Disease %in% c("PAN", "CAD healthy")) |> 
  select(all_of(overlapping_prots)) |> 
  cor(use = "complete.obs")

corrplot(corr_mat, 
         method = "square",
         addCoef.col = 'black',
         type = "lower", 
         order = "hclust", 
         hclust.method = "ward.D2",
         col = pal_corr,  
         tl.col = "black", 
         tl.srt = 0)

# Supplementary figure 5A
#ggsave(savepath("all_corr_ml.pdf"), h = 10, w = 10)

corr_mat_selected <- 
  cancer_data_wide |> 
  filter(Disease %in% c("PAN", "CAD healthy")) |> 
  select(all_of(top_overlapping_prots$Variable)) |> 
  cor(use = "complete.obs")

corrplot(corr_mat_selected, 
         method = "square",
         addCoef.col = 'black',
         type = "lower", 
         order = "hclust", 
         hclust.method = "ward.D2",
         col = pal_corr,  
         tl.col = "black", 
         tl.srt = 0)

# Figure 3D
ggsave(savepath("selected_corr_ml.pdf"), h = 10, w = 10)
```

## Importances

```{r}
overlapping_prots <- intersect(proteins_lasso, proteins_rf)

important_proteins_rf |> 
  select(Variable, Importance_rf = Importance) |> 
  left_join(important_proteins_da |> 
              mutate(Importance_lasso = abs(estimate)) |> 
              select(Variable = term, Importance_lasso), by = "Variable") |> 
  mutate(Importance_lasso = ifelse(is.na(Importance_lasso), 0, Importance_lasso),
         Importance_rf = ifelse(is.na(Importance_rf), 0, Importance_rf)) |> 
  mutate(Type = ifelse(Variable %in% overlapping_prots, "Overlapping", "Not overlapping")) |> 
  ggplot(aes(Importance_rf, Importance_lasso, color = Type)) +
  geom_point() +
  geom_text_repel(aes(label = Variable), show.legend = F) +
  scale_color_manual(values = c("grey70", "#A41E60")) +
  theme_hpa()

# Figure 3B
#ggsave(savepath("comparison_importances.pdf"), h = 6, w = 6)



important_proteins_rf |> 
  select(Variable, Importance_rf = Importance) |> 
  left_join(important_proteins_da |> 
              mutate(Importance_lasso = abs(estimate)) |> 
              select(Variable = term, Importance_lasso), by = "Variable") |> 
  mutate(Importance_lasso = ifelse(is.na(Importance_lasso), 0, Importance_lasso),
         Importance_rf = ifelse(is.na(Importance_rf), 0, Importance_rf)) |> 
  mutate(Type = ifelse(Variable %in% overlapping_prots, "Overlapping", "Not overlapping")) |> 
  mutate(Importance_rf = scales::rescale(Importance_rf, c(0,100)),
         Importance_lasso = scales::rescale(Importance_lasso, c(0,100))) |> 
  saveRDS("data/processed/ML_res/importances_combined.rds")


```


## Probabilities

```{r}

final_rf_fit |> 
  collect_predictions() |> 
  select(pancreatic_probability = `.pred_Pancreatic cancer`, Disease) |> 
  mutate(DAid = cancer_test$DAid,
         Model = "random forest") |> 
  bind_rows(final_glmnet_fit |> 
              collect_predictions() |> 
              select(pancreatic_probability = `.pred_Pancreatic cancer`, Disease) |> 
              mutate(DAid = cancer_test$DAid,
                     Model = "lasso")) |> 
  ggplot(aes(Model, pancreatic_probability, color = Model)) +
  geom_line(aes(group = DAid), color = "grey80", alpha = 0.3) +
  geom_quasirandom() +
 # geom_violin(fill = NA, color = "grey20", outlier.colour = NA) +
#  geom_boxplot(fill = NA, color = "grey20", outlier.colour = NA) +
  facet_wrap(~Disease) +
  scale_color_manual(values = c("#BBDB9B", "#A594BD")) +
  theme_hpa(angled = T)

ggsave(savepath("probabilties.pdf"), h = 5, w = 5)

```

# Signature against other cancers

### General

```{r}
all_wide_data <- 
  cancer_data_long |>
  select(DAid, Disease, Assay, NPX) |> 
  pivot_wider(names_from = "Assay", values_from = "NPX")

```

## glmnet

```{r}

pancreatic_against_cancers <-  
  function(cancer) {
    
    # Prepare data
    cancer_split_current <- 
      all_wide_data |>
      select(DAid, Disease, important_proteins_da$term) |> 
      filter(Disease == cancer) |> 
      initial_split(prop = 0.7, strata = Disease)
    
    training_dat <- 
      cancer_train |> 
      mutate(DAid = as.character(DAid)) |> 
      select(DAid, Disease, important_proteins_da$term) |> 
      filter(Disease == "Pancreatic cancer") |> 
      bind_rows(training(cancer_split_current))
    
    testing_dat <- 
      cancer_test |> 
      mutate(DAid = as.character(DAid)) |> 
      select(DAid, Disease, important_proteins_da$term) |> 
      filter(Disease == "Pancreatic cancer") |> 
      bind_rows(testing(cancer_split_current))
    
    cancer_split_custom <- make_splits(training_dat, testing_dat)
    
    cancer_recipe_2 <- 
      recipe(Disease ~ ., data = training_dat) |> 
      update_role(DAid, new_role = "id") |> 
      step_normalize(all_numeric()) |> 
      step_nzv(all_numeric()) |> 
      step_impute_knn(all_numeric()) #|> 
    #step_downsample(Disease)
    
    glmnet_specs_2 <- 
      logistic_reg() |> 
      set_mode("classification") |> 
      set_engine("glmnet") |> 
      set_args(penalty = tune(), 
               mixture = 1) 
    
    glmnet_wflow_2 <-
      workflow() |> 
      add_recipe(cancer_recipe_2) |> 
      add_model(glmnet_specs_2) 
    
    # Define glmnet grid
    set.seed(213)
    glmnet_grid_2 <-
      glmnet_wflow_2 |>
      extract_parameter_set_dials() |>
      grid_latin_hypercube(size = 30)
    
    # Define the resamples (CV)
    set.seed(213)
    cancer_rs_2 <- vfold_cv(training_dat, v = 10, strata = Disease)
    
    # Define the evaluation metrics (add brier)
    eval_metrics <- metric_set(roc_auc, accuracy, sensitivity, specificity)
    
    # Define control_grid
    set.seed(213)
    ctrl <- control_grid(save_pred = TRUE, parallel_over = "everything") # look at extract = identity
    
    # Glmnet grid search
    set.seed(213)
    glmnet_res_2 <-
      glmnet_wflow_2 |>
      tune_grid(
        resamples = cancer_rs_2,
        grid = glmnet_grid_2,
        control = ctrl,
        metrics = eval_metrics
      )
    
    autoplot(glmnet_res_2)
    
    best_glmnet_2 <- 
      select_best(glmnet_res_2, metric = "roc_auc") |> 
      select(-.config)
    
    glmnet_wflow_2 <- 
      glmnet_wflow_2 |>  
      finalize_workflow(best_glmnet_2)
    
    final_glmnet_fit_2 <- last_fit(glmnet_wflow_2, cancer_split_custom, metrics = eval_metrics) 
    
    performance <- 
      final_glmnet_fit_2 |> 
      collect_metrics() |> 
      select(-.config, -.estimator)
    
    glmnet_auc <- 
      final_glmnet_fit_2 |> 
      collect_metrics() |> 
      filter(.metric == "roc_auc") |> 
      pull(.estimate) |> 
      round(2)
    
    important_proteins_2 <- 
      final_glmnet_fit_2  |> 
      extract_fit_parsnip() %>%
      tidy() |> 
      filter(term != "(Intercept)") |> 
      arrange(-abs(estimate)) |> 
      filter(abs(estimate) > 0)
    
    predictions <- 
      final_glmnet_fit_2 |> 
      collect_predictions()
    
    cm <- 
      predictions |> 
      conf_mat(Disease, .pred_class) 
    
    roc <- 
      predictions |> 
      mutate(Disease = factor(Disease, levels = c("Pancreatic cancer",  cancer))) |> 
      roc_curve(truth = Disease, `.pred_Pancreatic cancer`) 
    
    
    return(list("penalty" = best_glmnet_2,
                "performance" = performance,
                "predictions" = predictions,
                "roc" = roc,
                "cm" = cm,
                "important_proteins" = important_proteins_2))
  }


# Run for all cancers
all_cancer_results <- 
  map(all_cancers_levels[-14], ~pancreatic_against_cancers(.)) |> 
  set_names(all_cancers_levels[-14])
```

### Importance scores

```{r}
# Extract importance inforamtion
imp_prots <- 
  map_df(all_cancers_levels[-14], function(cancer){
  
  res <- all_cancer_results[[cancer]]
  res$important_proteins |> 
    mutate(Cancer = cancer)
})
 
# imp_prots |> 
#   group_by(term) |> 
#   summarise(n = n_distinct(Cancer)) |> 
#   arrange(-n) |> 
#   filter(n > 1)
  
# Visualize
annotation_row <- 
  tibble(row = names(pal_cancer_controls),
         Cancer = names(pal_cancer_controls)) |> 
  column_to_rownames("row")

annotation_col <- 
  important_proteins_da |> 
  select(-penalty) |> 
  mutate(estimate = abs(estimate)) |> 
  rename(Estimate_original = estimate) |> 
  column_to_rownames("term")


imp_prots |> 
  left_join(cancers_mapping |> 
              select(-Cancer), by = c("Cancer" = "Cancer_code")) |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  select(term, estimate, Cancer = Cancer_short) |> 
  mutate(estimate = abs(estimate)) |> 
  spread(Cancer, estimate, fill = 0) |> 
  column_to_rownames("term") |> 
  t() |> 
  pheatmap(color = colorRampPalette(c("white", "darkred"))(100),
            annotation_row = annotation_row,
           annotation_col = annotation_col,
           annotation_colors = list("Cancer" = pal_cancer_controls),
           cellwidth = 5, cellheight = 5, fontsize = 5) |> as.ggplot()

# Figure 3G
ggsave(savepath("important_lasso_proteins_all_cancers.pdf"), h = 10, w = 5)
```

### ROC curves

```{r}
# Extract ROC curve information
all_roc <- 
  map_df(all_cancers_levels[-14], function(cancer){
  
  res <- all_cancer_results[[cancer]]
  auc <- 
    res$performance |> 
    filter(.metric == "roc_auc") |> 
    pull(.estimate)
  res$roc |> 
    mutate(Cancer = cancer,
           AUC = auc)
})

# Visualize results
pal_cancer_controls <- 
  cancers_mapping |> 
  filter(Cancer_code != "PAN") |> 
  group_by_all() |> 
  mutate(color = pal_all_cancers[[Cancer_code]]) |> 
  ungroup() |> 
  select(-Cancer_code, -Cancer) |> 
  deframe()

aucs <- 
  cancers_mapping |> 
  filter(Cancer_code != "PAN") |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  left_join(all_roc |> 
              distinct(Cancer, AUC), by = c("Cancer_code" = "Cancer")) |> 
    select(-Cancer_code,-Cancer) |> 
  mutate(y = c(0.65, 0.60, 0.55, 0.50, 
               0.45, 0.40, 0.35, 0.30,
               0.25, 0.20, 0.15, 0.10, 0.05),
         x = rep(0.75, 13))

all_roc |> 
  left_join(cancers_mapping, by = c("Cancer" = "Cancer_code")) |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  ggplot(aes(x = 1-specificity, y=sensitivity, group = Cancer_short, color = Cancer_short)) +
  geom_path(linewidth = 0.7, show.legend = F) +
  geom_segment(aes(x = 0, y = 0, xend = 1,  yend = 1), 
               colour='grey', linetype = 'dotdash', linewidth = 0.3) +
  geom_text(aes(x = 0.85, y = 0.70, label = "AUC"), fontface = "bold", color = "black", inherit.aes = F, size = 3) +
  geom_text(data = aucs, aes(x, y, label = Cancer_short), size = 3, show.legend = F, hjust = 1, fontface = "bold") +
  geom_text(data = aucs, aes(x + 0.10 , y, label = round(AUC,3)), size = 3, show.legend = F, color = "black") +
  theme_hpa() + 
  scale_color_manual(values = pal_cancer_controls) +
  coord_equal() +
  ggtitle("lasso")

# Figure 3F
#ggsave(savepath("lasso_other_cancers.pdf"), h = 5, w = 5)
```

### Probabilities

```{r}

```


## rf
```{r}

pancreatic_against_cancers_rf <-  
  function(cancer) {
    
    # Prepare data
    cancer_split_current <- 
      all_wide_data |>
      select(DAid, Disease, important_proteins_rf$Variable) |> 
      filter(Disease == cancer) |> 
      initial_split(prop = 0.7, strata = Disease)
    
    training_dat <- 
      cancer_train |> 
      mutate(DAid = as.character(DAid)) |> 
      select(DAid, Disease, important_proteins_rf$Variable) |> 
      filter(Disease == "Pancreatic cancer") |> 
      bind_rows(training(cancer_split_current))
    
    testing_dat <- 
      cancer_test |> 
      mutate(DAid = as.character(DAid)) |> 
      select(DAid, Disease, important_proteins_rf$Variable) |> 
      filter(Disease == "Pancreatic cancer") |> 
      bind_rows(testing(cancer_split_current))
    
    cancer_split_custom <- make_splits(training_dat, testing_dat)
    
    cancer_recipe_2 <- 
      recipe(Disease ~ ., data = training_dat) |> 
      update_role(DAid, new_role = "id") |> 
      step_normalize(all_numeric()) |> 
      step_nzv(all_numeric()) |> 
      step_impute_knn(all_numeric()) #|> 
    #step_downsample(Disease)
    
    cores <- parallel::detectCores()
    
    rf_specs_2 <-
      rand_forest() |>
      set_mode("classification") |>
      set_engine("ranger", importance = "impurity", num.threads = cores) |>
      set_args(mtry = tune(), min_n = tune(), trees = 1000) 
    
    # Set up random forest workflow
    rf_wflow_2 <-
      workflow() |> 
      add_recipe(cancer_recipe_2) |> 
      add_model(rf_specs_2) 
    
    # Define hyperparameter tuning grid
    set.seed(213) #test
    rf_grid_2 <-
      rf_wflow_2 |>
      extract_parameter_set_dials() |>
      update(mtry = mtry(c(0, 40))) |>
      grid_latin_hypercube(size = 15)
    
    # Define the resamples (CV)
    set.seed(213)
    cancer_rs_2 <- vfold_cv(training_dat, v = 10, strata = Disease)
    
    # Define the evaluation metrics (add brier)
    eval_metrics <- metric_set(roc_auc, accuracy, sensitivity, specificity)
    
    # Define control_grid
    set.seed(213)
    ctrl <- control_grid(save_pred = TRUE, parallel_over = "everything") # look at extract = identity
    
    # Hyperparameter tuning
    set.seed(213)
    rf_res_2 <-
      rf_wflow_2 |>
      tune_grid(
        resamples = cancer_rs_2,
        grid = rf_grid_2,
        control = ctrl,
        metrics = eval_metrics)
    
    best_rf_2 <- 
      select_best(rf_res_2, metric = "roc_auc") |> 
      select(-.config)
    
    rf_wflow_2 <- 
      rf_wflow_2 |>  
      finalize_workflow(best_rf_2)
    
    final_rf_fit_2 <- last_fit(rf_wflow_2, cancer_split_custom, metrics = eval_metrics) 
    
    performance <- 
      final_rf_fit_2 |> 
      collect_metrics() |> 
      select(-.config, -.estimator)
    
    glmnet_auc <- 
      final_rf_fit_2 |> 
      collect_metrics() |> 
      filter(.metric == "roc_auc") |> 
      pull(.estimate) |> 
      round(2)
    
    important_proteins_2 <- 
      final_rf_fit_2  |> 
      extract_fit_parsnip() |> 
      vip::vi() |> 
      arrange(-Importance) |> 
      filter(Importance > 0)
   
    predictions <- 
      final_rf_fit_2 |> 
      collect_predictions()
    
    cm <- 
      predictions |> 
      conf_mat(Disease, .pred_class) 
    
    roc <- 
      predictions |> 
      mutate(Disease = factor(Disease, levels = c("Pancreatic cancer",  cancer))) |> 
      roc_curve(truth = Disease, `.pred_Pancreatic cancer`) 
    
    
    return(list("penalty" = best_rf_2,
                "performance" = performance,
                "predictions" = predictions,
                "roc" = roc,
                "cm" = cm,
                "important_proteins" = important_proteins_2))
  }


# Run for all cancers
all_cancer_results_rf <- 
  map(all_cancers_levels[-14], ~pancreatic_against_cancers_rf(.)) |> 
  set_names(all_cancers_levels[-14])
```

### Importance scores

```{r}
# Extract importance inforamtion
imp_prots_rf <- 
  map_df(all_cancers_levels[-14], function(cancer){
  
  res <- all_cancer_results_rf[[cancer]]
  res$important_proteins |> 
    mutate(Cancer = cancer)
})
 
# imp_prots |> 
#   group_by(term) |> 
#   summarise(n = n_distinct(Cancer)) |> 
#   arrange(-n) |> 
#   filter(n > 1)
  
# Visualize
pal_cancer_controls

annotation_row <- 
  tibble(row = names(pal_cancer_controls),
         Cancer = names(pal_cancer_controls)) |> 
  column_to_rownames("row")

annotation_col <- 
  important_proteins_rf |> 
  rename(Importance_original = Importance) |> 
  column_to_rownames("Variable")

selected <- 
  imp_prots_rf |> 
  group_by(Variable) |> 
  summarize(imp = sum(Importance)) |> 
  arrange(-imp) |> 
  filter(imp > 3)

imp_prots_rf |> 
  filter(Variable %in% selected$Variable) |> 
  left_join(cancers_mapping |> 
              select(-Cancer), by = c("Cancer" = "Cancer_code")) |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  select(-Cancer) |> 
  rename(Cancer = Cancer_short) |> 
  select(Variable, Importance, Cancer) |> 
  spread(Cancer, Importance, fill = 0) |> 
  column_to_rownames("Variable") |> 
  t() |> 
  pheatmap(color = colorRampPalette(c("white", "darkred"))(100),
           annotation_row = annotation_row,
           annotation_col = annotation_col,
           annotation_colors = list("Cancer" = pal_cancer_controls),
           cellwidth = 5, cellheight = 5, fontsize = 5) |> as.ggplot()

# Figure 3G
ggsave(savepath("important_rf_proteins_all_cancers.pdf"), h = 10, w = 20)

```

### ROC curves

```{r}

# Extract ROC curve information
all_roc_rf <- 
  map_df(all_cancers_levels[-14], function(cancer){
  
  res <- all_cancer_results_rf[[cancer]]
  auc <- 
    res$performance |> 
    filter(.metric == "roc_auc") |> 
    pull(.estimate)
  res$roc |> 
    mutate(Cancer = cancer,
           AUC = auc)
})

# Visualize results
pal_cancer_controls <- 
  cancers_mapping |> 
  filter(Cancer_code != "PAN") |> 
  group_by_all() |> 
  mutate(color = pal_all_cancers[[Cancer_code]]) |> 
  ungroup() |> 
  select(-Cancer_code, -Cancer) |> 
  deframe()

aucs_rf <- 
  cancers_mapping |> 
  filter(Cancer_code != "PAN") |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  left_join(all_roc_rf |> 
              distinct(Cancer, AUC), by = c("Cancer_code" = "Cancer")) |> 
    select(-Cancer_code,-Cancer) |> 
  mutate(y = c(0.65, 0.60, 0.55, 0.50, 
               0.45, 0.40, 0.35, 0.30,
               0.25, 0.20, 0.15, 0.10, 0.05),
         x = rep(0.75, 13))

all_roc_rf |> 
  left_join(cancers_mapping, by = c("Cancer" = "Cancer_code")) |> 
  mutate(Cancer_short = factor(Cancer_short, levels = cancers_mapping$Cancer_short)) |> 
  ggplot(aes(x = 1-specificity, y=sensitivity, group = Cancer_short, color = Cancer_short)) +
  geom_path(linewidth = 0.7, show.legend = F) +
  geom_segment(aes(x = 0, y = 0, xend = 1,  yend = 1), 
               colour='grey', linetype = 'dotdash', linewidth = 0.3) +
  geom_text(aes(x = 0.85, y = 0.70, label = "AUC"), fontface = "bold", color = "black", inherit.aes = F, size = 3) +
  geom_text(data = aucs_rf, aes(x, y, label = Cancer_short), size = 3, show.legend = F, hjust = 1, fontface = "bold") +
  geom_text(data = aucs_rf, aes(x + 0.10 , y, label = round(AUC,3)), size = 3, show.legend = F, color = "black") +
  theme_hpa() + 
  scale_color_manual(values = pal_cancer_controls) +
  coord_equal() 

# Figure 3F
ggsave(savepath("rf_other_cancers.pdf"), h = 5, w = 5)
```

# Citations

```{r}
packageVersion("tidymodels")
packageVersion("themis")
packageVersion("colino")
packageVersion("vip")

devtools::session_info()
```
